---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# German North Sea

This package contains shapefiles from the **German North Sea**: <br>  

- [Natura2000](https://www.eea.europa.eu/data-and-maps/data/natura-13/natura-2000-spatial-data/natura-2000-shapefile-1)<br> 
- [Country shape](https://www.diva-gis.org/datadown)<br> 
- [Economic Exclusive zone](https://marineregions.org/gazetteer.php?p=details&id=5669)<br> 
- [ICES rectangles](https://gis.ices.dk/shapefiles/ICES_rectangles.zip)<br> 

The goal is provide easy access to shapefiles of the North Sea

All shapefiles are in CRS 3035.

## Installation

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("MiriamLL/GermanNorthSea")
```

```{r libreria}
library(GermanNorthSea)
```

# Data

### German Land

```{r}
German_land<-GermanNorthSea::German_land
```

```{r}
ggplot2::ggplot()+ ggplot2::geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

### German EEZ

```{r}
German_EEZ<-GermanNorthSea::German_EEZ
```

```{r}
ggplot2::ggplot()+ ggplot2::geom_sf(data = German_EEZ,
                                    colour = "red", fill= NA,alpha=0.9, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


### German Coast

```{r}
German_coast<-GermanNorthSea::German_coast
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = German_coast, 
                                    colour = "red", fill= '#bde0fe',alpha=0.9, lwd = 1,linetype="dashed")+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```



### German Natura 2000

```{r}
German_natura<-GermanNorthSea::German_natura
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

## ICES Rectangles

```{r}
German_ICES<-GermanNorthSea::German_ICES
```

```{r}
ggplot2::ggplot(German_ICES)+ 
  ggplot2::geom_sf(data = German_ICES, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::geom_sf_text(ggplot2::aes(label =ICESNAME),size=3,family="sans")+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```
# BATH

# Maps

```{r}
library(ggplot2)
library(ggspatial)
```

## German EEZ and land

```{r}
ggplot() +
  geom_sf(data = German_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+

  ggspatial::annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  theme(
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.background = element_rect(fill = '#bde0fe'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  ggplot2::coord_sf(xlim = c(3900000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

## German Natura 2000

```{r}
German_natura$Abbreviations<-c('SOR','BR','DB')
```

```{r}
ggplot(German_natura) +
  geom_sf(data = German_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+
  geom_sf_text(aes(label=Abbreviations),size=3,family="sans")+
  
  annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  theme(
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.background = element_rect(fill = '#bde0fe'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3900000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```




## ICES rectangles

```{r}
library(tidyverse)
```

```{r}
German_ICES<-GermanNorthSea::German_ICES
```

```{r}
ICES_rectangles<-German_ICES %>%
  filter(ICESNAME %in% c('40F3','40F4',
                         '39F4','39F5','39F6','39F7','39F8',
                         '38F5','38F6','38F7','38F8',
                         '37F5','37F6','37F7','37F8',
                         '36F6','36F7','36F8'))
```

```{r}
ggplot(ICES_rectangles)+ 
  geom_sf(data = ICES_rectangles, colour = "#14213d", fill= NA, alpha=0.2, lwd = 0.1)+
  geom_sf_text(aes(label=ICESNAME),size=3,family="sans")+

  geom_sf(data = German_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+

  ggspatial::annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  theme(
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.background = element_rect(fill = '#bde0fe'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  ggplot2::coord_sf(xlim = c(3900000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```



# Other CRS

For using other CRS, you can use the function **st_transform** from the package **sf**.

```{r}
library(sf)
```

```{r}
German_EEZ<-st_transform(German_EEZ, 4326)
German_coast<-st_transform(German_coast, 4326)
German_land<-st_transform(German_land, 4326)
German_natura<-st_transform(German_natura, 4326)
```

```{r}
ggplot()+
  geom_sf(data = German_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+

  ggspatial::annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  theme(
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.background = element_rect(fill = '#bde0fe'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

# Other sources

- [GEBCO](https://www.gebco.net/data_and_products/gridded_bathymetry_data/)

