---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# German North Sea

This package contains shapefiles from the **German North Sea**: <br>  

- [Natura2000](https://www.eea.europa.eu/data-and-maps/data/natura-13/natura-2000-spatial-data/natura-2000-shapefile-1)<br> 
- [Country shape](https://www.diva-gis.org/datadown)<br> 
- [Economic Exclusive zone](https://marineregions.org/gazetteer.php?p=details&id=5669)<br> 
- [ICES rectangles](https://gis.ices.dk/shapefiles/ICES_rectangles.zip)<br> 

The goal is provide easy access to shapefiles of the North Sea

All shapefiles are in CRS 3035.

## Installation

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("MiriamLL/GermanNorthSea")
```

```{r libreria}
library(GermanNorthSea)
```

# Data

## Base Maps
### German Land

To download go to: [DIVA-GIS](https://www.diva-gis.org/datadown)<br> 
Select the country and Administrative areas. 
```{r}
Germany<-GermanNorthSea::Germany
```

```{r}
ggplot2::ggplot()+ ggplot2::geom_sf(data = Germany, colour = 'black', fill = '#ffffbe')+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

### Europe
```{r}
German_land<-GermanNorthSea::German_land
```

```{r}
ggplot2::ggplot()+ ggplot2::geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

### German EEZ

To download: go to [EMODnet](https://www.emodnet-humanactivities.eu/view-data.php)>Select wind farms in the map.
Select Other Forms of Area Management/Designation>Click on the download icon on [Exclusive Economic Zone](https://www.emodnet-humanactivities.eu/search-results.php?dataname=Exclusive+Economic+Zone).
Select Download Data, Select the country, and the Sector, and the Link should be ready to download.

```{r}
German_EEZ<-GermanNorthSea::German_EEZ
```

```{r}
ggplot2::ggplot()+ ggplot2::geom_sf(data = German_EEZ,
                                    colour = "red", fill= NA,alpha=0.9, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


### German Coast

```{r}
German_coast<-GermanNorthSea::German_coast
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = German_coast, 
                                    colour = "red", fill= '#bde0fe',alpha=0.9, lwd = 1,linetype="dashed")+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```



### ICES Rectangles

Source: [ICES rectangles](https://gis.ices.dk/sf/index.html)<br> 
To download: go to link> click on **Quick Downloads**> select ICES Statistical Rectangles.

```{r}
German_ICES<-GermanNorthSea::German_ICES
```

```{r}
ggplot2::ggplot(German_ICES)+ 
  ggplot2::geom_sf(data = German_ICES, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::geom_sf_text(ggplot2::aes(label =ICESNAME),size=3,family="sans")+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

### German Natura 2000

To download: go to [GeoSeaPortal](https://www.geoseaportal.de/atomfeeds/Raumordnungsplan_AWZ_en.xml#download=1.)
The zip contains several shapefiles including **NatureConservation**

Other option is [eea](https://www.eea.europa.eu/data-and-maps/data/natura-13)

```{r}
German_natura<-GermanNorthSea::German_natura
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


Other option is [Europen Environment Agency](https://www.eea.europa.eu/data-and-maps/data/natura-14)
This file contains a more complete data from the natura 2000 sites.

For using the SCA (Special Conservation Areas) in Germany

```{r}
German_SCA<-GermanNorthSea::German_SCA
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = German_SCA, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

## Using both SPA and SCA

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::geom_sf(data = German_SCA, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


## Human activities
### Shipping lines

To download: go to [GeoSeaPortal](https://www.geoseaportal.de/atomfeeds/Raumordnungsplan_AWZ_en.xml#download=1.)
The zip contains several shapefiles including **Shipping**

```{r}
German_Shipping<-GermanNorthSea::German_Shipping
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = German_Shipping, colour = "blue", fill= 'blue',alpha=0.2, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


### Wind Farms


#### BfN
To download: go to [GeoSeaPortal](https://www.geoseaportal.de/atomfeeds/Raumordnungsplan_AWZ_en.xml#download=1.)
The zip contains several shapefiles including **'OffshoreWindEnergy'**

```{r}
German_OWF<-GermanNorthSea::German_OWF
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = German_OWF, colour = "#fe0001", fill= '#bde0fe',alpha=0.9, size=1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


#### EMODnet

To download: Go to [EMBO](https://www.emodnet-humanactivities.eu/view-data.php). Select [human activities](https://www.emodnet-humanactivities.eu/search-results.php?dataname=Wind+Farms+%28Polygons%29)

```{r}
OWF_EMODnet<-GermanNorthSea::OWF_EMODnet
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = OWF_EMODnet, colour = "#fe0001", fill= '#bde0fe',alpha=0.9, size=1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```






## Grids

This grids were created in R and are made available by this package.

To cite: Lerma, M (2023) German North Sea package. 

```{r}
library(ggplot2)
library(ggspatial)
```
### 5x5

```{r}
grid5x5_3035<-grid5x5_3035
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data=grid5x5_3035, colour = "black", fill= '#56C1D5', lwd = 0.5)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

### 10x10

```{r, eval=FALSE, echo=FALSE}
# Packages
library(tidyverse)
library(sf)

# Create limits
my_data<-data.frame(Longitude=c(3790000,4250000),
Latitude=c(3350000,3680000)
)

# Tranform to CRS
my_points<-my_data %>%
  st_as_sf(coords=c('Longitude','Latitude'),
           #crs=4326,
           crs=3035,
           remove=FALSE)

my_grid<-st_make_grid(my_points, 
                       c(10000, 10000), 
                       what = "polygons",
                       square = TRUE)
# Add id per grid
my_grid_sf = st_sf(my_grid) %>%
  dplyr::mutate(grid_id = 1:length(lengths(my_grid)))

# Check grid
ggplot()+
  geom_sf(data=my_grid, colour = "black", fill= '#56C1D5', lwd = 0.5)

# Asign CRS
grid10x10_3035<-my_grid_sf

# Export
#st_write(my_grid,"C:/Users/lerma/OneDrive/Documents/04Laboral/2023/01MarBird/2023/Overlaps/02Jun/1Data/03GIS/10x10grid_3035c.shp")

# rdata
save(grid10x10_3035, file = "grid10x10_3035.RData", version = 2)
```

```{r}
grid10x10_3035<-grid10x10_3035
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data=grid10x10_3035, colour = "black", fill= '#56C1D5', lwd = 0.5)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


```{r}
ggplot() +
  geom_sf(data = German_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  ggplot2::geom_sf(data=grid10x10_3035, colour = "black", fill= '#56C1D5', lwd = 0.5, alpha=0.1)+
    xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3900000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```



## Grid 10x10 EEZ

```{r, eval=FALSE, echo=FALSE}
# Package
library(GermanNorthSea)
library(tidyverse)
library(sf)
# Shapefiles
Grid_10km<-grid10x10_3035
Grid_10km_EEZ <- sf::st_intersection(Grid_10km, German_EEZ)
## With data
## Create data frame of grids with data by dropping geometry
Grid_names<-Grid_10km_EEZ %>% 
  st_drop_geometry()
## Select column with grid_id
Grid_names<-Grid_names %>%
  select(grid_id,OBJECTID)
## Grid inside EEZ
Grid_10km_keep<-merge(Grid_10km,Grid_names,by='grid_id',all=TRUE)
## Keep only with data
Grid_10km_keep<- Grid_10km_keep[stats::complete.cases(Grid_10km_keep$OBJECTID),]
# Asign name
grid10x10_EEZ<-Grid_10km_keep
# rdata
save(grid10x10_EEZ, file = "grid10x10_EEZ.RData", version = 2)

##devtools::document()
##devtools::check()
```

```{r}
grid10x10_EEZ<-grid10x10_EEZ
```

```{r}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data=grid10x10_EEZ, colour = "black", fill= '#56C1D5', lwd = 0.5)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


# Maps

```{r}
library(ggplot2)
library(ggspatial)
```

## Base Maps

```{r}
German_land<-GermanNorthSea::German_land
German_EEZ<-GermanNorthSea::German_EEZ
German_coast<-GermanNorthSea::German_coast
```

```{r}
ggplot() +
  geom_sf(data = German_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+

  annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  theme(
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.background = element_rect(fill = '#bde0fe'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3900000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

### ICES rectangles

```{r, message=FALSE}
library(tidyverse)
```

```{r}
German_ICES<-GermanNorthSea::German_ICES
```

```{r}
ICES_rectangles<-German_ICES %>%
  filter(ICESNAME %in% c('40F3','40F4',
                         '39F4','39F5','39F6','39F7','39F8',
                         '38F5','38F6','38F7','38F8',
                         '37F5','37F6','37F7','37F8',
                         '36F6','36F7','36F8'))
```

```{r}
ggplot(ICES_rectangles)+ 
  geom_sf(data = ICES_rectangles, colour = "#14213d", fill= NA, alpha=0.2, lwd = 0.1)+
  geom_sf_text(aes(label=ICESNAME),size=3,family="sans")+

  geom_sf(data = German_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+

  annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  theme(
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.background = element_rect(fill = '#bde0fe'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3900000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

## Protected areas

```{r}
German_natura<-GermanNorthSea::German_natura
```

```{r}
German_natura$Abbreviations<-c('SOR','BR','DB')
```

```{r}
ggplot(German_natura) +
  geom_sf(data = German_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+
  geom_sf_text(aes(label=Abbreviations),size=3,family="sans")+
  
  annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  theme(
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.background = element_rect(fill = '#bde0fe'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3900000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```



## Human activities
### Wind Farms

```{r}
OWF_EMODnet<-GermanNorthSea::OWF_EMODnet
```

```{r}
ggplot(OWF_EMODnet)+ 
  geom_sf(data = German_EEZ,colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast,colour = "red", fill= '#bde0fe',alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+
  geom_sf(data = OWF_EMODnet, aes(fill= STATUS), colour = "grey",alpha=0.9, size=0.5)+
  scale_fill_manual(values = c("yellow", "blue", "green"))+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme(legend.position = 'bottom')
```

### Shipping lines

```{r}
German_Shipping<-GermanNorthSea::German_Shipping
```

```{r}
ggplot(German_Shipping)+ 
  geom_sf(data = German_EEZ,colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast,colour = "red", fill= '#bde0fe',alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+
  geom_sf(data = German_Shipping, aes(fill= Type), colour = "grey",alpha=0.9, size=0.5)+
  scale_fill_manual(values = c("blue", "deepskyblue", "turquoise"))+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme(legend.position = 'bottom')
```

### Shipping intensity




# Other CRS

For using other CRS, you can use the function **st_transform** from the package **sf**.

```{r}
library(sf)
```

```{r}
library(GermanNorthSea)
library(ggspatial)
```

```{r}
German_EEZ<-st_transform(German_EEZ, 4326)
German_coast<-st_transform(German_coast, 4326)
German_land<-st_transform(German_land, 4326)
German_natura<-st_transform(German_natura, 4326)
```

```{r, fig.height=8, fig.width=8}
ggplot()+
  geom_sf(data = German_EEZ, colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "red", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+

  ggspatial::annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  theme(
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.background = element_rect(fill = '#bde0fe'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```

#### GEBCO Bathymetry

To download [GEBCO: General Bathymetry Chart of the Oceans](https://www.gebco.net/data_and_products/gridded_bathymetry_data/#area)

Go to: Download data for user-defined areas
Use the application.
Add your coordinates here I use 1 to 10 and 50 to 60.
Add to basket and download.

https://download.gebco.net/


```{r, include=FALSE}
library(here)
here()
this_folder<-paste0(here(),"/data")
list.files(this_folder)
this_file<-paste0(this_folder,"/gebco_2024_n60.0_s50.0_w1.0_e9.0.tif")
this_file
```

Use the [terra package](https://datacarpentry.org/r-raster-vector-geospatial/01-raster-structure.html)

```{r}
library(terra)
```

```{r}
Bath_file<-rast(this_file)
```

```{r}
Bath_dataframe <- as.data.frame(Bath_file, xy = TRUE)
```

# Alternatively

```{r}
Bath_dataframe<-GermanNorthSea::German_bath
```

```{r}
Bath_dataframe_sub <-Bath_dataframe  %>%
  filter(x > 2 & x < 10)%>%
  filter(y > 52 & y < 57)%>%
  rename(Bathymetry=3) %>%
  filter(Bathymetry < 10)%>%
  mutate(Bathymetry = as.numeric(Bathymetry))
```

```{r}
library(ggplot2)
```

```{r}
ggplot() +
  geom_raster(data = Bath_dataframe_sub , aes(x = x, y = y, fill = Bathymetry)) +
  scale_fill_viridis_c(option = "mako")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))

```


```{r}
German_land<-st_transform(German_land, 4326)
```

```{r}
ggplot() +
  geom_raster(data = Bath_dataframe_sub , aes(x = x, y = y, fill = Bathymetry)) +
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  scale_fill_viridis_c(option = "mako")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))

```

Do not forget to cite:

If the data sets are used in a presentation or publication then we ask that you acknowledge the source.This should be of the form:
GEBCO Compilation Group (2024) GEBCO 2024 Grid (doi:10.5285/1c44ce99-0a0d-5f4f-e063-7086abc0ea0f)

# Other sources

- [Marine Spatial Planning](https://www.bsh.de/EN/TOPICS/Offshore/Maritime_spatial_planning/maritime_spatial_planning_node.html)

- [Shipping intensity](https://www.emodnet-humanactivities.eu/download-data.php)
- [Marine Traffic](https://www.marinetraffic.com/en/ais/home/centerx:8.6/centery:53.9/zoom:9)

- [Offshore wind farms](https://map.4coffshore.com/offshorewind/)
- [Windkraftanlagen](https://commons.wikimedia.org/wiki/File:Karte_Offshore-Windkraftanlagen_in_der_Deutschen_Bucht.png)

- [GEBCO](https://www.gebco.net/data_and_products/gridded_bathymetry_data/)
- [NOAA-GRIDDAP-GLOBAL](https://erddap.ifremer.fr/erddap/griddap/index.html?page=1&itemsPerPage=1000)
- [COPERNICUS](https://resources.marine.copernicus.eu/products)
